---
title: "HW04"
author: "63130500041 Tawan Muadmuenwai"
output: rmarkdown::github_document
---

# Contents

1. [Define the questions](#define-the-questions)
2. [Explore the dataset](#explore-the-dataset)
3. [Data Transformation and Data Cleaning](#data-transformation-and-data-cleaning)
4. [Exploratory Data Analysis](#exploratory-data-analysis)
5. [Data Visualization](#data-visualization)
6. [Functions not taught in the class](#functions-not-taught-in-the-class)


# Define the questions

1. What are the top ten most revenue-generating sub-categories?
2. What are the top ten most ordered products in 2017?
3. What are the average time to ship in each year?
4. What are the top ten states with the highest number of customers?
5. What are the top ten highest-spending customers?
6. Is the sales increasing or decreasing over time?


# Explore the dataset

## Import libraries

```{r lib, message=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
```


## Import the dataset

```{r import, message=FALSE}
superstore <-
  read_csv("https://raw.githubusercontent.com/safesit23/INT214-Statistics/main/datasets/superstore_sales.csv")
```


## Explore the dataset

```{r explore}
glimpse(superstore)
```

This dataset has 9800 observations and 18 variables. Each one of them is self-explanatory.

- Row ID is a `numeric` variable.
- Order ID is a `character` variable.
- Order Date is a `character` variable.
- Ship Date is a `character` variable.
- Ship Mode is a `character` variable.
- Customer ID is a `character` variable.
- Customer Name is a `character` variable.
- Segment is a `character` variable.
- Country is a `character` variable.
- City is a `character` variable.
- State is a `character` variable.
- Postal Code is a `numeric` variable.
- Region is a `character` variable.
- Product ID is a `character` variable.
- Category is a `character` variable.
- Sub-Category is a `character` variable.
- Product Name is a `character` variable.
- Sales is a `numeric` variable.

### What should be done?

- Order Date and Ship Date should be converted to date objects.
- Ship Mode, Customer ID, Segment, Country, City, State, Postal Code, Region, Product ID, Category, Sub-Category should be converted to factor variables.


# Data Transformation and Data Cleaning

## 1. Convert the date-formatted columns to date objects
```{r date_conversion}
superstore <-
  superstore %>%
  mutate(`Order Date` = as_date(`Order Date`, format = "%d/%m/%Y")) %>%
  mutate(`Ship Date` = as_date(`Ship Date`, format = "%d/%m/%Y"))
```


## 2. Convert categorical columns to factors
```{r categorical_conversion}
superstore <-
  superstore %>%
    mutate(
      across(c(`Ship Mode`, `Customer ID`, Segment, Country, City, State, `Postal Code`, Region, `Product ID`, Category, `Sub-Category`), factor))
```



# Exploratory Data Analysis

## 1. What are the top ten most revenue-generating sub-categories?

```{r top_ten_sub_categories}
superstore %>%
  group_by(`Sub-Category`) %>%
  summarise(
    Sales = sum(`Sales`)) %>%
  arrange(desc(Sales)) %>%
  head(10)
```

**Summary**

The top ten most revenue-generating sub-categories are:

- Phones with a total sales of $327,782
- Chairs with a total sales of $322,823
- Storage with a total sales of $219,343
- Tables with a total sales of $202,811
- Binders with a total sales of $200,029
- Machines with a total sales of $189,239
- Accessories with a total sales of $164,187
- Copiers with a total sales of $146,248
- Bookcases with a total sales of $113,813
- Appliances with a total sales of $104,618


## 2. What are the top ten most ordered products in 2017?

```{r top_ten_ordered_products}
superstore %>%
  filter(year(`Order Date`) == 2017) %>%
  group_by(`Product Name`) %>%
  summarise(OrderCount = n()) %>%
  arrange(desc(OrderCount)) %>%
  head(10)
```

**Summary**

The top ten most ordered product in 2017 are:

- "Staple envelope" with a total order count of 11
- "Easy-staple paper" with a total order count of 10
- "Chromcraft Round Conference Tables" with a total order count of 9
- "Staples" with a total order count of 9
- "XtraLife ClearVue Slant-D Ring Binder, White, 3\"" with a total order count of 7
- "Acco Perma 4000 Stacking Storage Drawers" with a total order count of 6
- "Avery Non-Stick Binders" with a total order count of 6
- "Hon Deluxe Fabric Upholstered Stacking Chairs, Rounded Back" with a total order count of 6
- "Ibico EB-19 Dual Function Manual Binding System" with a total order count of 6
- "Premium Transparent Presentation Covers by GBC" with a total order count of 6


## 3. What are the average time to ship in each year?

```{r average_time_to_ship}
superstore %>%
  mutate(Year = year(`Ship Date`)) %>%
  mutate(DaysToShip = `Ship Date` - `Order Date`) %>%
  group_by(Year) %>%
  summarise(
    AverageDaysToShip = mean(`DaysToShip`) %>% round(2)) %>%
  arrange(Year)
```

**Summary**

The average time to ship in each year are:

- 2015 with an average time to ship of 3.98 days
- 2016 with an average time to ship of 4.04 days
- 2017 with an average time to ship of 3.93 days
- 2018 with an average time to ship of 3.91 days
- 2019 with an average time to ship of 4.93 days


## 4. What are the top ten states with the highest number of customers?

```{r state_with_most_customers}
superstore %>%
  group_by(State) %>%
  distinct(`Customer ID`) %>%
  summarise(
    CustomerCount = n()) %>%
  arrange(desc(CustomerCount)) %>%
  head(10)
```

**Summary**

The top ten states with the highest number of customers are:

- California with a total customer count of 570
- New York with a total customer count of 409
- Texas with a total customer count of 367
- Pennsylvania with a total customer count of 255
- Illinois with a total customer count of 231
- Washington with a total customer count of 223
- Ohio with a total customer count of 196
- Florida with a total customer count of 178
- North Carolina with a total customer count of 120
- Virginia with a total customer count of 107


## 5. What are the top ten highest-spending customers?

```{r top_ten_highest_spending_customer}
superstore %>%
  group_by(`Customer ID`) %>%
  mutate(
    AmountSpent = sum(Sales)) %>%
  distinct(`Customer ID`, `Customer Name`, AmountSpent) %>%
  arrange(desc(AmountSpent)) %>%
  head(10)
```

**Summary**

The top ten highest-spending customer are:

- Sean Miller with a total amount spent of $25,043
- Tamara Chand with a total amount spent of $19,052
- Raymond Buch with a total amount spent of $15,117
- Tom Ashbowl with a total amount spent of $14,596
- Adrian Barton with a total amount spent of $14,474
- Ken Lonsdale with a total amount spent of $14,175
- Sanjit Chand with a total amount spent of $14,142
- Hunter Lopez with a total amount spent of $12,873
- Sanjit Engle with a total amount spent of $12,209
- Christopher Conant with a total amount spent of $12,129


## 6. Is the sales increasing or decreasing over time?

```{r sales_increasing_or_decreasing}
sales_by_year <-
  superstore %>%
  mutate(
    Year = year(`Order Date`)) %>%
  group_by(Year) %>%
  summarise(
    Sales = sum(Sales)) %>%
  arrange(Year)

cor(sales_by_year$Year, sales_by_year$Sales)
```

**Summary**

The sales are increasing over time because the years and sales have a strong positive correlation.



# Data Visualization

## 1. What are the top ten highest-spending customers?

```{r viz_top_ten_ordered_products}
superstore %>%
  group_by(`Customer ID`) %>%
  mutate(
    AmountSpent = sum(Sales)) %>%
  distinct(`Customer ID`, `Customer Name`, AmountSpent) %>%
  arrange(desc(AmountSpent)) %>%
  head(10) %>%
  ggplot(aes(x = reorder(`Customer Name`, -AmountSpent), y = AmountSpent)) +
    geom_col(fill = "#b00335") +
    scale_y_continuous(labels = scales::dollar_format()) +
    labs(x = "Customer Name", y = "Amount Spent", title = "Top 10 Highest-Spending Customers") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**Summary**

The sales drop for a bit in 2016 but then go back up tremendously in 2017, continuing to go up in 2018, which means that overall, the sales are increasing over time.


## 2. What are the sales trends for each month of each year?

```{r viz_sales_trend_setup, message=FALSE}
sales_by_month <- superstore %>%
  mutate(
    Year = year(`Order Date`),
    Month = month(`Order Date`)) %>%
  group_by(Year, Month) %>%
  summarise(
    Sales = sum(Sales)) %>%
  arrange(Year, Month)
```

```{r viz_sales_trend}
sales_by_month %>%
  ggplot(aes(x = Month, y = Sales, color = as.factor(Year))) +
    geom_line(alpha = 0.8) +
    geom_point() +
    labs(color = "Year", title = "Monthly Sales Revenue from 2015-2018") +
    scale_y_continuous(labels = scales::dollar_format()) +
    scale_x_continuous(labels = c(month.abb[1:12]), breaks = c(1:12)) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "top")
```

**Summary**

In 2015, the sales started off with a relatively low amount of sales, but then increased dramatically in March and dropped again in April. The sales increased dramatically in September and then dropped again in October, followed by a strong increase in November and dipped for a bit in December.

2016 is similar to 2015, but the changes in sales are less dramatic.

In 2017, the sales started off lower than in 2016, but then increased dramatically in March and gradually dropped for a bit in August. The sales increased dramatically in September and kept increasing to the end of the year.

2018 started off with highest sales among 4 years, but then dropped dramatically in Feburary. The sales went through the same pattern as in 2017, but the changes are more dramatic and hit the peak amount of sales per one month in November.

Overall, each year starts off with a low sales amount, but then gradually increases in March, occasionally drops and then increases again between Q2 and Q3, ending with a peak amount of sales in Q4.



# Functions not taught in the class

## 1. `as_date()`

Convert date-formatted text to a date. A drop-in replacement for `as.Date()`.

### Example

```{r as_date}
"27/05/2042" %>% as_date(format = "%d/%m/%Y")
```


## 2. `year()`

Extracts the year from a date.

### Example

```{r year}
"27/05/2042" %>% as_date(format = "%d/%m/%Y") %>% year()
```


## 3. `month()`

Extracts the month from a date.

### Example

```{r month}
"27/05/2042" %>% as_date(format = "%d/%m/%Y") %>% month()
```


## 4. `across()`

Used to apply a function across multiple columns. Can be used inside `mutate()`.

### Example

```{r across_setup}
data <- data.frame(
  Id = 1:10,
  Letter = LETTERS[1:10])

data %>% glimpse()
```

Supposed we want to convert both columns Id (`numeric`) and Letter (`character`) to `factor`

```{r across}
data %>%
  mutate(
    across(c(Id, Letter), factor)) %>%
  glimpse()
```
